---
title: "52414 Lab Encryption Guide"
author: "52414"
date: "4/25/2018"
output: html_document
---
A short guide on file encryption using public key and `sodium`  

First you need to install/load the necessary Libraries.  
Both `Sodium` and `repmis` are in CRAN (and thus can be installed via `install.packages()` ).
However, `cypher` must be installed from github using the command that starts with `remotes::`
```{r}
library("sodium")
library("remotes")
library("repmis") #see line 17
# To install cypher:
#remotes::install_github("ropensci/cyphr", upgrade = FALSE)
library("cyphr")
```

Loading the key object from github:  
Go to [Public_Key](https://github.com/DataScienceHU/DataAnalysisR/blob/master/guides/pub_cyphr_52414?raw=true) and the public key should be downloaded to your computer. 

```{r}
path.to.public.key <- "C:/Users/Alpha/Documents/university/yearc/simesterB/DataAnalysiswithR/labs/lab2/pub_cyphr_52414"
load(path.to.public.key) #loading the downloaded public key into R
```


Encrypt your lab rmd file:
Now you are ready to encrypt your lab1 rmd file.  
Specify the path of your rmd file and the path you want to save your encripted rmd file in the following script:

```{r}
print(getwd()) #this can help you locate your rmd file
rmd.path.name.input <- "C:/Users/Alpha/Documents/university/yearc/simesterB/DataAnalysiswithR/labs/lab2/lab_2.rmd"
rmd.path.name.output <- "C:/Users/Alpha/Documents/university/yearc/simesterB/DataAnalysiswithR/labs/lab2/lab2_group14_crypt.rmd"
# Below is the actual encryption command:
cyphr::encrypt_file(rmd.path.name.input, pub_cyphr , rmd.path.name.output)
```

Last step: Upload your encypted RMD file to github in the following link:
`https://github.com/DataScienceHU/DataAnalysisR/tree/master/Lab1/Lab1_submissions/`
