---
title: "Lab 2"
date: "Due 06/6/2018"
output: html_document
editor_options: 
chunk_output_type: inline
---

GROUP Number: ____fill-in____  
Name of 1st group member: Gioia Blayer  
Email of 1st group member: gioia.blayer@mail.huji.ac.il  
Name of 2st group member: Efrat Aviad 
Email of 2st group member: efrat.aviad@mail.huji.ac.il

**Instructions:**   
1. Please rename your group lab_2 rmd file as - `Lab2_group#.rmd`.   
** The # should be replaced by your group number.  
2. Fill in your code in the relevant code execution boxes.  
3. Please avoid plagiarism - do not share your answers with other groups.  
4. Use the encryption guide (see moodle) to encrypt your final rmd file and name it `Lab2_group#_crypt`   
5. Upload the encrypted file to the following link:  
** `https://github.com/DataScienceHU/DataAnalysisR/tree/master/Lab2/Lab2_submission/`

# Lab 2

Vaccines have helped save millions of lives. In the 19th century, before herd immunization was achieved through vaccination programs, deaths from infectious diseases, like smallpox and polio, were common. However, today, despite all the scientific evidence for their importance, vaccination programs have become somewhat controversial.

The controversy started with a [paper](http://www.thelancet.com/journals/lancet/article/PIIS0140-6736(97)11096-0/abstract) published in 1988 and lead by [Andrew Wakefield](https://en.wikipedia.org/wiki/Andrew_Wakefield) claiming there was a link between the administration of the measles, mumps and rubella (MMR) vaccine, and the appearance of autism and bowel disease. Despite much science contradicting this finding, sensationalists media reports and fear mongering from conspiracy theorists, led parts of the public to believe that vaccines were harmful. Some parents stopped vaccinating their children. This dangerous practice can be potentially disastrous given that the Center for Disease Control (CDC) estimates that vaccinations will prevent more than 21 million hospitalizations and 732,000 deaths among children born in the last 20 years (see [Benefits from Immunization during the Vaccines for Children Program Era United States, 1994-2013, MMWR](https://www.cdc.gov/mmwr/preview/mmwrhtml/mm6316a4.htm)). 

Effective communication of data is a strong antidote to misinformation and fear mongering. In this homework you are going to prepare a report to have ready in case you need to help a family member, friend or acquaintance that is not aware of the positive impact vaccines have had for public health.

The data used for these plots were collected, organized and distributed by the [Tycho Project](http://www.tycho.pitt.edu/). They include weekly reported counts data for seven diseases from 1928 to 2011, from all fifty states.


```{r, required libraries, include=FALSE}
library(ggplot2)
library(tidyverse)
library(RColorBrewer)
```

## Part 1

0. Load the `diseases` data from Moodle or Github  

    **Solution:** 
```{, load data}
us_contagious_diseases <- read.csv("diseases.csv")
```
```{r, load data}
admin <- "Lee" #  "Lee" # or 
if (admin=="Yuval") {
  us_contagious_diseases_name <- "/Users/yuvalb/Dropbox/52414/Lectures_2018/course_data/disease_data/diseases"
} else if (admin == "Lee"){
  us_contagious_diseases_name <- "~/Dropbox (Personal)/School/52414/Lectures_2018/course_data/disease_data/diseases"
} else {
  print("Admin name not specified correctly")
}

us_contagious_diseases <- read.csv(file = us_contagious_diseases_name, stringsAsFactors = FALSE)
```


1. Use the `us_contagious_disease` and `dplyr` tools to create a data.frame called `dat` that stores only the Measles data (with all the other features). Please add  a per 100,000 people rate, and remove Alaska and Hawaii from the data since they only became states in the late 50s. Note that there is a `weeks_reporting` column. Take that into account when computing the rate.  

    **Solution:** 
```{r}
library(dplyr)
dat <- us_contagious_diseases %>% filter((disease=="Measles" & state!="Alaska" & state!= "Hawaii"))

##dat$rate <- (dat$count/dat$weeks_reporting)*(dat$population/100000)  
dat$rate <- (52*dat$count/dat$weeks_reporting)*(dat$population/100000)  # LEe's addition       
dat$rate[is.nan(dat$rate)] <- 0 
```
**Note that we changed the Nan values into zeros because the Nan was created by dividing in zeros. The rate itself is zero, because in those states they didn't have any measles at this year at all.**

```{,Lee's Comments}
The rate calculation is incorrect... you should multiply the LHS by 52 to normalize all the data to a year - your current numbers are in terms of 1 week. 
```


2. Plot the Measles disease rates per year for California. Find out when the Measles vaccine was introduced and add a vertical line to the plot to show this year.  

    **Solution:** 
```{r}
dat.cali <- dat %>% filter(state=="California")

library(reshape2)
ggplot(dat.cali, aes(x=year,y=rate))+
  geom_line(color="blue")+
  geom_vline(xintercept = 1963,color="red")+
  geom_text(aes(label="Measles Vaccine Introduction", x = 1964, y = 170000),color = "red", hjust = -.02)+
  ggtitle("California's Measles Rate per Year")
  
```

**After a short research online, we discovered that the measles faccination was introduced in 1963.**  

3. Does the pattern hold for other states? Use boxplots to get an idea of the distribution of rates for each year, and see if the pattern holds across states.  
a. Generate the plot with scale_y_continuous 
b. Generate the same plot but now with scale_y_sqrt (breaks = seq(0, 100, 4)^2))  
Which one is better and why? 

    **Solution:**   
    
```{r}

#a
ggplot(data=dat, aes(x=abbreviate(state,2), y=rate, frame = year, size=population))+
  geom_boxplot()+
  scale_y_continuous(breaks = seq(0, 350000, 10000))+
  theme(legend.position = "none",text = element_text(size=7),
  axis.text.x = element_text(angle=45, hjust=1))+
  xlab("State")+
  ggtitle("US Measles Rates per State")
  
#b
ggplot(data=dat, aes(x=abbreviate(state,2), y=rate, frame = year, size=population))+
  geom_boxplot()+
  scale_y_sqrt (breaks = seq(0, 1000, 20)^2)+
  theme(legend.position = "none",text = element_text(size=7),
  axis.text.x = element_text(angle=45, hjust=1))+
  xlab("State")+
  ylab("Rate in Squared Root")+
  ggtitle("US Measles Rates in squared scale per State")
  
```

**In the first graph, we have outliers, which do not enable us to infer conclusions about the distribution of rates for each state. The second graph is better because it's easier to decipher for the reader. In addition, we need to remember in order to make any conclusions, that not all the data is taken into account, and that the rate is in squared root.**
```{, Lee's comments}
We asked for each year  - basically to aggregate the states together so it would be more comparable to the last question... 
```


4. One problem with the boxplot is that it does not let us see state-specific trends.  Plot the Measles disease rates per year for all states. Find out when the Measles vaccine was introduced and add a vertical line to the plot to show this year. Add the US average to the plot. Hint: Note there are missing values in the data.  

    **Solution:**  
   
```{r}

year_count <- dat %>% group_by(year, state) 
years <- 1928:2003
US.mean <- tapply(dat$rate,dat$year,FUN = mean)
US.dat <- as.vector(US.mean)
av.US <- data.frame(years,US.dat)

library(directlabels)

ggplot() +
    geom_line(data = year_count, aes(x = year, y = rate , color = state))+
    #scale_y_continuous(breaks = seq(0, 300000, 50000))+
  scale_y_sqrt("Cases per 100,000", breaks = seq(0, 100, 4)^2) + ##Lee's addition
    theme(legend.position="none")+
    geom_vline(xintercept = 1963, size=1, col="red")+ 
    geom_line(data = av.US, aes(x = av.US$years, y = av.US$US.dat)
              ,size = 1,color="black")+
  geom_text(aes(label = "US Mean", x = 1995, colour = av.US, y = 20000), hjust = -.02)+
  geom_text(aes(label = "1963 - Measles Introduction Year", x = 1964, y = 270000),color = "red", hjust = -.02)+
  ggtitle("Measles disease rates per year for all states")
            
```

## Part 2

Here we'll start analyzing the results we collected in class regarding the display of visual information. 

5. Download and import the "Cleveland Experiment.csv" and "correct.Rdata" files from moodle. 

```{}
experiment.data <- read.csv("Cleveland_Experiment.csv")
load("C:/Users/Gioia/Documents/Hebrew University/Materie Curricolari/Second Year/Statistica/Semester B/Data Analysis with R/Labs/Lab 2/correct.rdata")
experiment.data <- experiment.data[-1,-1]

```
```{r}
admin <- "Lee" #  "Lee" # or 
if (admin=="Yuval") {
  experiment <- "/Users/yuvalb/Dropbox/52414/Lectures_2018/labs/lab_2/Cleveland_Experiment.csv"
  correct <- "/Users/yuvalb/Dropbox/52414/Lectures_2018/course_data/labs/lab_2/correct.rdata"
} else if (admin == "Lee"){
  experiment <- "~/Dropbox (Personal)/School/52414/Lectures_2018/labs/lab_2/Cleveland_Experiment.csv"
  correct <- "~/Dropbox (Personal)/School/52414/Lectures_2018/labs/lab_2/correct.rdata"
} else {
  print("Admin name not specified correctly")
}

experiment.data <- read.csv(file = experiment, stringsAsFactors = FALSE)
load(correct)
```


6. Look at the distribution of guesses of Question 2 of the experiment.  
a. Compare the correct difference vs the reported (experiment data)
b. Compare it to a normal distribution with a histogram and with the qq-plot. 
c. Compare to the normal distribution using a density plot. 
d. Would you approximate this distribution using the normal? 

    Solution:

```{r}
library(ggplot2)
#a
ggplot(experiment.data,aes(Question.2))+
  geom_histogram(binwidth = 0.03)+
  geom_vline(xintercept = correct[2],col="green",size=2)+
  xlab("Question 2 Answers")+
  ylab("Frequency")+
  ggtitle("Question 2 Answer's distribution versus the Correct answer")

```

```{r}
#b.1
mean_answer <- mean(experiment.data$Question.2)
sd_answer <- sd(experiment.data$Question.2)

ggplot(experiment.data,aes(Question.2))+ 
  geom_histogram(binwidth = 0.03)+
  stat_function(fun=dnorm,args = list(mean_answer,sd_answer),col="red")+
  xlab("Question 2 Answers")+
  ylab("Frequency")+
  ggtitle("Question 2 Answer's distribution compared to Normal distribution")

```
```{, Lee's comments}
Where is teh normal vs. the experiment data? Go over solutions... 
```

```{r}
#b.2
Question.2 <-experiment.data$Question.2
q2 <- (Question.2-mean(Question.2))/sd(Question.2)
ggplot(data=experiment.data,aes(sample=q2))+
  geom_qq()+geom_abline(size=1,col="red")+
  ggtitle("Quantile Comparison between Question 2 answers and Normal distribution")

```

```{, Lee's comments}
I only see a line... , no qq dots.. 
```

```{r}
#c
ggplot(experiment.data,aes(Question.2))+
  geom_density(fill="green",alpha=.3)+
  stat_function(fun=dnorm,args = list(mean_answer,sd_answer),size = 1,color = "darkgreen")+
  xlab("Question 2 Answer")+
  ggtitle("Question 2 density versus normal distribution")
```
#d 
```{, Lee's comments}
This is the density of the experiment, what about the density of the normal?
```

**No, we wouldn't approximate this distribution using the normal. The distribution of the data is not simmetrical, and the tails are considerably big for the normal distribution. Moreover, the number of samples is considerably small in our opinion to be compared to the normal distribution.** 


7. Perform the same analysis (all the above) for question 5 of the experiment.  
How are the two compare?  

    Solution:

```{r}
ggplot(experiment.data,aes(Question.5))+
  geom_histogram(binwidth = 0.03)+
  geom_vline(xintercept = correct[5],col="blue",size=2)+
  ggtitle("Question 5 Answer's distribution versus the Correct answer")


```

```{r}
#b.1
mean_answer.5 <- mean(experiment.data$Question.5)
sd_answer.5 <- sd(experiment.data$Question.5)

ggplot(experiment.data,aes(Question.5))+ 
  geom_histogram(binwidth = 0.03)+
  stat_function(fun=dnorm,args = list(mean_answer.5,sd_answer.5),col="blue")+
  ggtitle("Question 5 Answer's distribution compared to Normal distribution")


```

```{r}
#b.2
Question.5 <-experiment.data$Question.5
q5 <- (Question.5-mean(Question.5))/sd(Question.5)
ggplot(data=experiment.data,aes(sample=q5))+
  geom_qq()+geom_abline(size=1,col="blue")+
  ggtitle("Quantile Comparison between Question 5 answers and Normal distribution")

```

```{r}
#c
ggplot(experiment.data,aes(Question.5))+
  geom_density(fill="#56B4E9",alpha=.5)+
  stat_function(fun=dnorm,args = list(mean_answer.5,sd_answer.5),size = 1,color="blue")+
    ggtitle("Question 5 density versus normal distribution")

```

**We can see that question 5 doesn't fit with the normal distribution, strongly than question 2. It still presents a right tail and it's not simmetrical, and the qqplot shows that the quantiles don't fit. Since the sample is small, the outliers have a more significant impact on the distribution, thus deforming the graph and deviating the reader's conclusions.**


8. In one figure graph the boxplots for each experiment question. Hint - use the base R function `stack`. 
a. Do you think the output figure is satisfying in terms of explaining the data? 
b. Use scale_y_continuous to trim some of the outliers to make the figure more readable. In addition, add a line to each boxplot to symbolize the `true` mean.  Hint - one way of adding the true lines is by adding another set of boxplots with zero variation centered around `correct`.  

    Solutions:
```{r}
colnames(experiment.data) <- c(1:20)
quest.vec <- stack(experiment.data)

ggplot(quest.vec,aes(x=quest.vec$ind,y=quest.vec$values, col=quest.vec$ind))+
  geom_boxplot()+
  xlab("Question Number")+
  ylab("Question Values")+
  theme(legend.position="none")+
  ggtitle("Boxplot of all the questions' distribution ")
```

**a. No, since we did not omit the outliers that included those people who answered non-reasonable values, when the correct answer was on a scale 0-0.6. Omitting the outliers would enable us to infer the correct conclusion about the distribution of the data. In the current chart, the outliers distort the image.**

```{r}
cor <- rep(correct,each=24)
colnames(experiment.data) <- c(1:20)
cor.mat <- matrix(cor,nrow=24,ncol=20)
cor.data <- as.data.frame(cor.mat)
colnames(cor.data) <- c(1:20)
cor.vec <- stack(cor.data)
quest.vec.1 <- stack(experiment.data)

ggplot(NULL, aes(ind,values))+
  geom_boxplot(data = quest.vec.1)+
  geom_boxplot(data = cor.vec, col="red")+
  scale_y_continuous(limits = c(0,1))+
  geom_text(aes(label = "Red is the correct difference", x = 11, y = 1),color = "red", hjust = -.02)+
  xlab("Question Number")+
  ggtitle("Boxplot of all the questions' distribution without outliers")

```
