---
title: "Lab 2"
date: "Due 06/6/2018"
output: html_document
editor_options: 
chunk_output_type: inline
---

GROUP Number: 12  
Name of 1st group member: Eitan Rosenfelder  
Email of 1st group member: Eitan.Rosenfelder@mail.huji.ac.il 
Name of 2st group member: Yoel Fraenkel 
Email of 2st group member: Yoel.Fraenkel@mail.huji.ac.il

**Instructions:**   
1. Please rename your group lab_2 rmd file as - `Lab2_group#.rmd`.   
** The # should be replaced by your group number.  
2. Fill in your code in the relevant code execution boxes.  
3. Please avoid plagiarism - do not share your answers with other groups.  
4. Use the encryption guide (see moodle) to encrypt your final rmd file and name it `Lab2_group#_crypt`   
5. Upload the encrypted file to the following link:  
** `https://github.com/DataScienceHU/DataAnalysisR/tree/master/Lab2/Lab2_submission/`

# Lab 2

Vaccines have helped save millions of lives. In the 19th century, before herd immunization was achieved through vaccination programs, deaths from infectious diseases, like smallpox and polio, were common. However, today, despite all the scientific evidence for their importance, vaccination programs have become somewhat controversial.

The controversy started with a [paper](http://www.thelancet.com/journals/lancet/article/PIIS0140-6736(97)11096-0/abstract) published in 1988 and lead by [Andrew Wakefield](https://en.wikipedia.org/wiki/Andrew_Wakefield) claiming there was a link between the administration of the measles, mumps and rubella (MMR) vaccine, and the appearance of autism and bowel disease. Despite much science contradicting this finding, sensationalists media reports and fear mongering from conspiracy theorists, led parts of the public to believe that vaccines were harmful. Some parents stopped vaccinating their children. This dangerous practice can be potentially disastrous given that the Center for Disease Control (CDC) estimates that vaccinations will prevent more than 21 million hospitalizations and 732,000 deaths among children born in the last 20 years (see [Benefits from Immunization during the Vaccines for Children Program Era — United States, 1994-2013, MMWR](https://www.cdc.gov/mmwr/preview/mmwrhtml/mm6316a4.htm)). 

Effective communication of data is a strong antidote to misinformation and fear mongering. In this homework you are going to prepare a report to have ready in case you need to help a family member, friend or acquaintance that is not aware of the positive impact vaccines have had for public health.

The data used for these plots were collected, organized and distributed by the [Tycho Project](http://www.tycho.pitt.edu/). They include weekly reported counts data for seven diseases from 1928 to 2011, from all fifty states.


```{r, required libraries, include=FALSE}
library(ggplot2)
library(tidyverse)
library(RColorBrewer)
library(dplyr)
library(utils)
```

## Part 1

0. Load the `diseases` data from Moodle or Github  

    **Solution:** 
```{, load data}

us_contagious_diseases <- read.csv("diseases")
```
```{r, load data}
admin <- "Lee" #  "Lee" # or 
if (admin=="Yuval") {
  us_contagious_diseases_name <- "/Users/yuvalb/Dropbox/52414/Lectures_2018/course_data/disease_data/diseases"
} else if (admin == "Lee"){
  us_contagious_diseases_name <- "~/Dropbox (Personal)/School/52414/Lectures_2018/course_data/disease_data/diseases"
} else {
  print("Admin name not specified correctly")
}

us_contagious_diseases <- read.csv(file = us_contagious_diseases_name, stringsAsFactors = FALSE)
```
```{r}
admin <- "Lee" #  "Lee" # or 
if (admin=="Yuval") {
  experiment <- "/Users/yuvalb/Dropbox/52414/Lectures_2018/labs/lab_2/Cleveland_Experiment.csv"
  correct <- "/Users/yuvalb/Dropbox/52414/Lectures_2018/course_data/labs/lab_2/correct.rdata"
} else if (admin == "Lee"){
  experiment <- "~/Dropbox (Personal)/School/52414/Lectures_2018/labs/lab_2/Cleveland_Experiment.csv"
  correct <- "~/Dropbox (Personal)/School/52414/Lectures_2018/labs/lab_2/correct.rdata"
} else {
  print("Admin name not specified correctly")
}

experiment.data <- read.csv(file = experiment, stringsAsFactors = FALSE)
load(correct)
```


1. Use the `us_contagious_disease` and `dplyr` tools to create a data.frame called `dat` that stores only the Measles data (with all the other features). Please add  a per 100,000 people rate, and remove Alaska and Hawaii from the data since they only became states in the late 50s. Note that there is a `weeks_reporting` column. Take that into account when computing the rate. 

    **Solution:** 
```{r}
dat<- filter(us_contagious_diseases, disease=="Measles" & state!="Alaska" & state!="Hawaii" & weeks_reporting > 0)
dat$rate100K <- (dat$count / dat$population * 100000 *365/7/dat$weeks_reporting)

"here we added a column which calculates the rate of the measles disease for each 100,000 people based on the weekly countings done"
```

2. Plot the Measles disease rates per year for California. Find out when the Measles vaccine was introduced and add a vertical line to the plot to show this year.  

    **Solution:** 
```{r}
filter(dat, state == "California") %>%
  ggplot( aes(x =year, y = rate100K))+
  geom_point()+
  geom_vline(xintercept = 1963)+
  ggtitle("Measles per year in California")

"we ploted here the rate of the spreading of the disease in california over the years, we marked the year 1963 when the vaccine was introduced. as we can see there was a huge drop in the rate of the dissease shortly after."

```


3. Does the pattern hold for other states? Use boxplots to get an idea of the distribution of rates for each year, and see if the pattern holds across states.  
a. Generate the plot with scale_y_continuous
b. Generate the same plot but now with scale_y_sqrt (breaks = seq(0, 100, 4)^2))  
Which one is better and why?

    **Solution:**   
    
```{r}
#a
ggplot(dat, aes(x= year, y= rate100K, group = year))+
  geom_boxplot()+
  ggtitle("plot of boxplots- US Measles per year")

#b
ggplot(dat, aes(x= year, y= rate100K, group = year))+
  geom_boxplot()+
  scale_y_sqrt (breaks = seq(0, 100, 4)^2)+
  ggtitle("scale_y_sqrt- US Measles per year")

"we can see that the same pattern continues in all the stats together to. we can see two plots which are built by diffrent frequencies, each one is better for different uses."
  
```



4. One problem with the boxplot is that it does not let us see state-specific trends.  Plot the Measles disease rates per year for all states. Find out when the Measles vaccine was introduced and add a vertical line to the plot to show this year. Add the US average to the plot. Hint: Note there are missing values in the data.  

    **Solution:**  
   
```{r}
us_avg <- aggregate(dat$rate100K, by=list(dat$year), FUN=mean)


dat %>%
  ggplot( aes(x = year, y = rate100K))+
    geom_point( color= "green")+
    geom_vline(xintercept = 1963)+ 
    geom_point(data= us_avg, aes(Group.1,x))+
    ggtitle("Measles per year- average in black")

"here we can see the rate for all states on a yearly basis in green, with a black dot which showes the yearly average."
```
```{}
Choice of colors is poor - see solution. 
```


## Part 2

Here we'll start analyzing the results we collected in class regarding the display of visual information. 

5. Download and import the "Cleveland Experiment.csv" and "correct.Rdata" files from moodle. 

```{, load data}

experiment.data <- read.csv("C:\\Users\\yoel fraenkel\\Desktop\\Cleveland_Experiment.csv")
 load("C:/Users/yoel fraenkel/Desktop/lab 2/correct.rdata")
```


6. Look at the distribution of guesses of Question 2 of the experiment.  
a. Compare the correct difference vs the reported (experiment data)
b. Compare it to a normal distribution with a histogram and with the qq-plot.  
c. Compare to the normal distribution using a density plot. 
d. Would you approximate this distribution using the normal? 

"we can see the gap beatween the guesses for the second question, by the students to the real diffrence at the first histogram. the second histogram is comparing the answers given distributed by a histogram and a normal distribution. the third graph shows qqplot. the fourth shows a plot based on a density plot."
   
    Solution:

```{r}
tmp_q2<- experiment.data$Question.2
q2<- as.numeric(tmp_q2[-1:-2])

ggplot(as.data.frame(q2),aes(q2))+
  geom_bar(breaks = 10, density = 10)+
  geom_vline(xintercept = correct[2])+
  ggtitle("correct difference vs the reported")
```

```{, Lee's comments}
Too many bins - increase binwidth. See solutions. 
```


```{r}
ggplot(as.data.frame(q2),aes(q2))+
  geom_bar(breaks = 10, density = 10)+
  ggtitle("histogram Compared to  a normal distribution")+
  stat_function(fun = dnorm, n = 1000, args = list(mean = mean(q2), sd = sd(q2)))

```

```{r}
ggplot(experiment.data[complete.cases(experiment.data), ], aes(sample = Question.2)) +
  stat_qq()+
  ggtitle("qqplot")

```
```{,Lee's comments}
45 degree line missing from qq plot. 
```


```{r}

ggplot(as.data.frame(q2),aes(q2))+
  geom_density()+
  ggtitle("density Compared to  a normal distribution")+
  stat_function(fun = dnorm, n = 1000, args = list(mean = mean(q2), sd = sd(q2)))

```
```{,Lee's comments}
Labels missing . 
```


```{r}
"this distribution looks approximately like a normal distribution, but actually its not close enough for most of the statistical useings"

```

7. Perform the same analysis (all the above) for question 5 of the experiment.  
How are the two compare?  

    Solution:


```{r}
tmp_q5<- experiment.data$Question.5
q5<- as.numeric(tmp_q5[-1:-2])

ggplot(as.data.frame(q5),aes(q5))+
  geom_bar(breaks = 10, density = 10)+
  geom_vline(xintercept = correct[2])+
  ggtitle("correct difference vs the reported")
```

```{r}
ggplot(as.data.frame(q5),aes(q5))+
  geom_bar(breaks = 10, density = 10)+
  ggtitle("histogram Compared to  a normal distribution")+
  stat_function(fun = dnorm, n = 1000, args = list(mean = mean(q5), sd = sd(q5)))

```
```{r}
ggplot(experiment.data[complete.cases(experiment.data), ], aes(sample = Question.5)) +
  stat_qq()+
  ggtitle("qqplot")

```
```{r}

ggplot(as.data.frame(q5),aes(q5))+
  geom_density()+
  ggtitle("density Compared to  a normal distribution")+
  stat_function(fun = dnorm, n = 1000, args = list(mean = mean(q5), sd = sd(q5)))
```
```{r}
"this distribution is definitely not even close to a normal distribution"

```

8. In one figure graph the boxplots for each experiment question. Hint - use the base R function `stack`. 
a. Do you think the output figure is satisfying in terms of explaining the data? 
b. Use scale_y_continuous to trim some of the outliers to make the figure more readable. In addition, add a line to each boxplot to symbolize the `true` mean.  Hint - one way of adding the true lines is by adding another set of boxplots with zero variation centered around `correct`.  

    Solutions:
```{r}
data_stack<-stack(experiment.data[-1:-2,-1]) 
ggplot(data_stack,aes(ind,as.factor(values),group = ind)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  ggtitle("boxplot for each experiment")

#we think this figure is satisfying in terms of explaining the data, it's impossible to understand it if you are not a statistics student""
```

```{r}
data_stack$the_real_min<- rep(correct,each=23)
data_stack %>% ggplot(aes(x = as.factor(ind),y = as.numeric(values)))+
  geom_boxplot() + 
  geom_boxplot(aes(x=as.factor(ind), y= the_real_min),linetype = 4, col = "blue")+
  scale_y_continuous(limits =c(0,1))+
   theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  ggtitle("boxplot for each experiment with mean in blue")

```



